A Responsible Machine Learning Workflow\\\normalsize{With Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2020-02-02 article-information-2019.tex
2/1 revisions to fit tables

As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on debunking the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN results. The bulk of the post-hoc explanation and discrimination testing results for the simulated are left to Appendix \ref{a_sec:res_sim}.

@@

Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Accuracy is measured at the best F1 threshold for each model. Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. $g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results. Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Accuracy is measured at the best F1 threshold for each model.

\begin{table}[H]
	\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
	\centering
	\begin{tabular}{ccccc}
		\toprule
		\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
		\midrule
		$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
		$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
		$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
		$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
		\bottomrule
		\label{tab:sim_acc}
	\end{tabular}
\end{table}

@@

\begin{table}[H]
\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccccc}
\toprule
\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
\midrule	
$g^\text{GBM}$		& 0.795	& 0.828	& 0.252	& 0.276 \\
$g^\text{MGBM}$	& 0.765	& 0.814	& 0.259	& 0.278 \\
$g^\text{ANN}$		& 0.865	& 0.871	& 0.231	& 0.262 \\
$g^\text{XNN}$		& 0.869	& 0.868	& 0.233	& 0.263 \\
\bottomrule
\label{tab:mort_acc}
\end{tabular}
\end{table}

>>>>

As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on challenging the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN interpretability results. The bulk of the post-hoc explanation and discrimination testing results for the simulated are left to Appendix \ref{a_sec:res_sim}.

@@

Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Accuracy is measured at the best F1 threshold for each model. Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. $g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results. Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Measures that require a probability cutoff are taken at the best F1 threshold for each model.

\begin{table}[H]
	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Arrows indicate the direction of improvement for each measure.}
	\centering
	\begin{tabular}{ccccccccc}
		\toprule
		\textbf{Model} & 
		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ & 
		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
		\midrule
		$g^\text{GBM}$		& 0.775	& 0.857	& & 0.474 &	& 0.394 & & \\
		$g^\text{MGBM}$		& 0.763	& 0.846	& & 0.498 &	& 0.405 & & \\
		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 &	0.525 & 0.398 & 0.858 & \\
		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 &	0.528 & 0.397 & 0.867 & \\
		\bottomrule
		\label{tab:sim_acc}
	\end{tabular}
\end{table}

@@

\begin{table}[H]
	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Arrows indicate the direction of improvement for each measure.}
\centering
\begin{tabular}{ccccccccc}
	\toprule
	\textbf{Model} & 
	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ & 
	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
	\midrule	
$g^\text{GBM}$		& 0.795	& 0.828	& & 0.252 &	& 0.276 & & \\
$g^\text{MGBM}$		& 0.765	& 0.814	& & 0.259 &	& 0.278 & & \\
$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 &	0.418 & 0.262 & 0.624 & \\
$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 &	0.409 & 0.263 & 0.593 & \\
\bottomrule
\label{tab:mort_acc}
\end{tabular}
\end{table}
